### 2021-08-25 analyses.RmD
- setting up repo - changed initial submission already strongly

---
title: "Who Needs Privacy? Exploring the relationship between personality and need for privacy"
subtitle: "Analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 90)
```

```{r r-setup, include=F}
# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# define packages
packages <- c("corrplot", "devtools", "GGally", "ggplot2", "haven", "kableExtra", "knitr", 
              "lavaan", "magrittr", "mice", "MVN", "PerFit", "psych", "quanteda.textstats",
              "pwr", "semPlot", "semTools", "tidyverse", "td")

# load packages
lapply(packages, library, character.only = TRUE)

# readability function
readability = function(data){
  reading_estimates <- c("Flesch.Kincaid", "FOG", "Coleman.Liau.short", "Dale.Chall.PSK")
  textstat_readability(data, measure = reading_estimates) %>% 
    as.data.frame %>% 
    mutate(average = rowMeans(select(., -document)))
}
```

```{r data-wrangling}
d <- read_spss("data/Need+for+Privacy_August+16%2C+2021_15.33.sav")

# define vars and labels
vars_pers_fac <- c("HEX_HOH_SIN", "HEX_HOH_FAI", "HEX_HOH_GRE", "HEX_HOH_MOD", 
                   "HEX_EMO_FEA", "HEX_EMO_ANX", "HEX_EMO_DEP", "HEX_EMO_SEN", 
                   "HEX_EXT_SSE", "HEX_EXT_BOL", "HEX_EXT_SOC", "HEX_EXT_LIV", 
                   "HEX_AGR_FOR", "HEX_AGR_GEN", "HEX_AGR_FLX", "HEX_AGR_PAT", 
                   "HEX_CNS_ORG", "HEX_CNS_DIL", "HEX_CNS_PER", "HEX_CNS_PRU", 
                   "HEX_OPN_AES", "HEX_OPN_INQ", "HEX_OPN_CRE", "HEX_OPN_UNC", 
                   "HEX_ALT")
vars_pers_fac_txt <- c("Sincerity", "Fairness", "Greed avoidance", "Modesty", 
                       "Fearfulness", "Anxiety", "Dependence", "Sentimentality", 
                       "Social Self-Esteem", "Social Boldness", "Sociability", "Liveliness",
                       "Forgiveness", "Gentleness", "Flexibility", "Patience", 
                       "Organization", "Diligence", "Perfectionism", "Prudence", 
                       "Aesthetic Appreciation", "Inquisitiveness", "Creativeness", "Unconventionality", 
                       "Altruism")

vars_pers_dim <- c("HEX_HOH", "HEX_EMO", "HEX_EXT", "HEX_AGR", "HEX_CNS", "HEX_OPN")
vars_pers_dim_txt <- c("Honesty humility", "Emotionality", "Extraversion", 
                       "Agreeableness", "Conscientiousness", "Openness")

vars_pri <- c("NFP_INF", "NFP_PSY", "NFP_PHY", "NFP_CNC", "NFP_SPA", 
              "NFP_GOV", "NFP_INT", "NFP_ANO", "NFP_COM", "NFP_GEN")
vars_pri_txt <- c("Informational", "Psychological", "Physical", "Concealment", "Personal Space", 
                  "Government", "Interpersonal", "Anonymity", "Companies", "General")

vars_ses <- c("age", "gender", "ethnicity", "relationship", "college", "income", "conservatism")

```

# Items

All items were answered on a 7-point scale with the following options:

"To what extent do you agree or disagree with the following statements?"

(-3) Strongly Disagree, (-2) Disagree, (-1) Slightly Disagree, (0) Neutral, (1) Slightly Agree (2) Agree, (3) Strongly Agree

## Personality
### Honesty Humility
#### Sincerity

 - If I want something from a person I dislike, I will act very nicely toward that person in order to get it.
 - I wouldn't use flattery to get a raise or promotion at work, even if I thought it would succeed.
 - If I want something from someone, I will laugh at that person's worst jokes.
 - I wouldn't pretend to like someone just to get that person to do favors for me.

```{r}
model <- "HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04"
```

#### Fairness

 - If I knew that I could never get caught, I would be willing to steal a million dollars.
 - I would be tempted to buy stolen property if I were financially tight.
 - I would never accept a bribe, even if it were very large.
 - I’d be tempted to use counterfeit money, if I were sure I could get away with it.

```{r}
model <- "HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04"
```

#### Greed Avoidance

 - Having a lot of money is not especially important to me.
 - I would like to live in a very expensive, high-class neighborhood.
 - I would like to be seen driving around in a very expensive car.
 - I would get a lot of pleasure from owning expensive luxury goods.

```{r}
model <- "HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04"
```

#### Modesty

 - I am an ordinary person who is no better than others.
 - I wouldn’t want people to treat me as though I were superior to them.
 - I think that I am entitled to more respect than the average person is.
 - I want people to know that I am an important person of high status.

```{r}
model <- "HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04"
```

### Emotionality
#### Fearfulness

 - I would feel afraid if I had to travel in bad weather conditions.
 - I don’t mind doing jobs that involve dangerous work.
 - When it comes to physical danger, I am very fearful.
 - Even in an emergency I wouldn't feel like panicking.

```{r}
model <- "HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04"
```

#### Anxiety

 - I sometimes can't help worrying about little things.
 - I worry a lot less than most people do.
 - I rarely, if ever, have trouble sleeping due to stress or anxiety.
 - I get very anxious when waiting to hear about an important decision.

```{r}
model <- "HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04"
```

#### Dependence
 - When I suffer from a painful experience, I need someone to make me feel comfortable.
 - I can handle difficult situations without needing emotional support from anyone else.
 - Whenever I feel worried about something, I want to share my concern with another person.
 - I rarely discuss my problems with other people.

```{r}
model <- "HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04"
```

#### Sentimentality

 - I feel like crying when I see other people crying.
 - When someone I know well is unhappy, I can almost feel that person's pain myself.
 - I feel strong emotions when someone close to me is going away for a long time.
 - I remain unemotional even in situations where most people get very sentimental.

```{r}
model <- "HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04"
```

### Extraversion
#### Social Self-Esteem

 - I feel reasonably satisfied with myself overall.
 - I think that most people like some aspects of my personality.
 - I feel that I am an unpopular person.
 - Even in an emergency I wouldn't feel like panicking.

```{r}
model <- "HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04"
```

#### Social Boldness

 - I rarely express my opinions in group meetings.
 - In social situations, I'm usually the one who makes the first move.
 - When I'm in a group of people, I'm often the one who speaks on behalf of the group.
 - I tend to feel quite self-conscious when speaking in front of a group of people.

```{r}
model <- "HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04"
```

#### Sociability

 - I avoid making "small talk" with people.
 - I enjoy having lots of people around to talk with.
 - I prefer jobs that involve active social interaction to those that involve working alone.
 - The first thing that I always do in a new place is to make friends.

```{r}
model <- "HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04"
```

#### Liveliness

 - I am energetic nearly all the time.
 - On most days, I feel cheerful and optimistic.
 - People often tell me that I should try to cheer up.
 - Most people are more upbeat and dynamic than I generally am.

```{r}
model <- "HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04"
```

### Agreeableness
#### Forgiveness

 - I rarely hold a grudge, even against people who have badly wronged me.
 - My attitude toward people who have treated me badly is "forgive and forget".
 - If someone has cheated me once, I will always feel suspicious of that person.
 - I find it hard to fully forgive someone who has done something mean to me.

```{r}
model <- "HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04"
```

#### Gentleness

 - People sometimes tell me that I am too critical of others.
 - I generally accept people’s faults without complaining about them.
 - I tend to be lenient in judging other people.
 - Even when people make a lot of mistakes, I rarely say anything negative.

```{r}
model <- "HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04"
```

#### Flexibility

 - People sometimes tell me that I'm too stubborn.
 - I am usually quite flexible in my opinions when people disagree with me.
 - When people tell me that I’m wrong, my first reaction is to argue with them.
 - I find it hard to compromise with people when I really think I’m right.

```{r}
model <- "HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04"
```

#### Patience

 - People think of me as someone who has a quick temper.
 - I rarely feel anger, even when people treat me quite badly.
 - Most people tend to get angry more quickly than I do.
 - I find it hard to keep my temper when people insult me.

```{r}
model <- "HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04"
```

### Conscientiousness
#### Organization

 - I clean my office or home quite frequently.
 - I plan ahead and organize things, to avoid scrambling at the last minute.
 - People often joke with me about the messiness of my room or desk.
 - When working, I sometimes have difficulties due to being disorganized.

```{r}
model <- "HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04"
```

#### Diligence

 - When working, I often set ambitious goals for myself.
 - I often push myself very hard when trying to achieve a goal.
 - Often when I set a goal, I end up quitting without having reached it.
 - I do only the minimum amount of work needed to get by.

```{r}
model <- "HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04"
```

#### Perfectionism

 - I often check my work over repeatedly to find any mistakes.
 - When working on something, I don't pay much attention to small details.
 - I always try to be accurate in my work, even at the expense of time.
 - People often call me a perfectionist.

```{r}
model <- "HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04"
```

#### Prudence

 - I make decisions based on the feeling of the moment rather than on careful thought.
 - I make a lot of mistakes because I don't think before I act.
 - I don’t allow my impulses to govern my behavior.
 - I prefer to do whatever comes to mind, rather than stick to a plan.

```{r}
model <- "HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04"
```

### Openness
#### Aesthetic Appreciation

 - I would be quite bored by a visit to an art gallery.
 - I wouldn't spend my time reading a book of poetry.
 - If I had the opportunity, I would like to attend a classical music concert.
 - Sometimes I like to just watch the wind as it blows through the trees.

```{r}
model <- "HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04"
```

#### Inquisitiveness

 - I'm interested in learning about the history and politics of other countries.
 - I enjoy looking at maps of different places.
 - I would be very bored by a book about the history of science and technology.
 - I’ve never really enjoyed looking through an encyclopedia.

```{r}
model <- "HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04"
```

#### Creativeness

 - I would like a job that requires following a routine rather than being creative.
 - I would enjoy creating a work of art, such as a novel, a song, or a painting.
 - People have often told me that I have a good imagination.
 - I don't think of myself as the artistic or creative type.

```{r}
model <- "HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04"
```

#### Unconventionality

 - I think that paying attention to radical ideas is a waste of time.
 - I like people who have unconventional views.
 - I think of myself as a somewhat eccentric person.
 - I find it boring to discuss philosophy.

```{r}
model <- "HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04"
```

### Altruism

 - I have sympathy for people who are less fortunate than I am.
 - I try to give generously to those in need.
 - It wouldn’t bother me to harm someone I didn’t like.
 - People see me as a hard-hearted person.

```{r}
model <- "HEX_ALT =~ HEX_ALT_01 + HEX_ALT_02 + HEX_ALT_03 + HEX_ALT_04"
```

## Need for privacy
### Informational

 - I would prefer that little is known about me. 
 - In general, I prefer to remain incognito. 
 - I don’t want my personal data to be publicly accessible. 
 - Not everyone has to know everything about me.

```{r}
model <- "NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04"
```

### Psychological

 - I don’t like to talk about personal issues unless others do it first.
 - There are a lot of things about me that I don’t like to talk about with others. 
 - I feel uncomfortable when others tell me private things about their lives. 
 - It is hard for me to talk about my feelings and inner thoughts.

```{r}
model <- "NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04"
```

### Physiological

 - I don’t like it when strangers come too close to me. 
 - I feel uncomfortable when others enter my flat or room unannounced. 
 - I don’t like to stand in a dense crowd of people. 
 - I don’t like to sit next to a stranger on the tram or bus.

```{r}
model <- "NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04"
```

### Privacy from other People

 - My need for privacy is so strong that it prevents me from using social media actively. 
 - I don’t feel I need to tell my friends all my secrets.
 - I feel I need to protect my privacy from other people.
 - I prefer it when others don’t know much about me.

```{r}
model <- "NFP_INT =~ NFP_INT_01 + NFP_INT_02 + NFP_INT_03 + NFP_INT_04"
```

#### Validation

### Privacy from Government

 - I need government agencies to respect my privacy, even if that makes catching criminals harder.,
 - I don’t want the government to gather data about me, even if that makes it harder to spend tax income efficiently.,
 - I don’t want government agencies to monitor my communication, even if that prevents terrorist attacks.,
 - I feel I need to protect my privacy from government agencies.

```{r}
model <- "NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04"
```

#### Validation

### Privacy need anonymity

 - I need to be able to use a fake account on social network sites to preserve my privacy. 
 - I feel I need to avoid places with video surveillance.
 - I prefer not to carry my drivers license or ID with me all the time to preserve my privacy.
 - I need to be able to surf the Internet anonymously. (new)

```{r}
model <- "NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04"
```

#### Validation

### Privacy from Companies

 - I’m willing to pay more for products or services so that companies don’t have to sell my data.
 - I’m happy to give up a little bit of privacy, so that I can use certain apps or services for free.
 - I don’t want companies to gather information about me, even if that reduces the quality of their services.
 - I feel I need to protect my privacy from companies.

```{r}
model <- "NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04"
```

#### Validation

### General Need for privacy

 - I need a lot of privacy.
 - Privacy is very important to me.
 - I think a lot about how I can protect my privacy.
 - I need a lot of time for myself.

```{r}
model <- "NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04"
```

#### Validation

# Power Analyses

To estimate sample size, we conducted a rudimentary power analysis. That is, we planned for a two-sided correlation of r = .1. We adopted a balanced alpha and beta error approach. Because we were in exploration mode, we aimed for a high statistical power and relaxed alpha. Ideally, we would have wanted to set an error rate of 5%; however, this was outside of our budget. Instead, we opted for an error rate of 10%.

```{r}
sesoi <- .1

# a priori power analyses
pwr.r.test(r = sesoi, sig.level = .05, power = .95) ## too expensive
```

A 5% error rate would imply collecting 1.293 participants, which was outside of our budget.

```{r}
pwr.r.test(r = sesoi, sig.level = .1, power = .9)
```

A 10% error rate would imply collecting 853 participants, which was within our budget.

```{r}
power <- pwr.r.test(r = sesoi, sig.level = .05, power = .8)
print(power)
power_n <- power$n
```

# Results
## Facets

```{r}
model_fac <- "
HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04
HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04
HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04
HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04
HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04
HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04
HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04
HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04
HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04
HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04
HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04
HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04
HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04
HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04
HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04
HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04
HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04
HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04
HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04
HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04
HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04
HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04
HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04
HEX_ALT =~ HEX_ALT_01 + HEX_ALT_02 + HEX_ALT_03 + HEX_ALT_04
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_CNC =~ NFP_CNC_01 + NFP_CNC_02 + NFP_CNC_03 + NFP_CNC_04
NFP_SPA =~ NFP_SPA_01 + NFP_SPA_02 + NFP_SPA_03 + NFP_SPA_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_INT =~ NFP_INT_01 + NFP_INT_02 + NFP_INT_03 + NFP_INT_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
age =~ AGE
gender =~ GEN
ethnicity =~ ETH
relationship =~ REL_
college =~ COL_
income =~ INC
conservatism =~ CON
"
fit_fac <- sem(model_fac, d, estimator = "ML", fixed.x = TRUE)
summary(fit_fac, standardized = TRUE)

d_fac <- 
  standardizedsolution(fit_fac, ci = TRUE, level = .90) %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_pers_dim, rhs %in% vars_pri)

fig_fac <- ggplot(d_fac, aes(x = est.std, y = lhs)) +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper)) +
  facet_wrap(vars(rhs), nrow = 1)

# tab_results <- data.frame(
#   tab_results_raw[1 : length(vars_pers_fac), ]$std.all,
#   tab_results_raw[(1 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 2), ]$std.all,
#   tab_results_raw[(2 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 3), ]$std.all,
#   tab_results_raw[(3 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 4), ]$std.all,
#   tab_results_raw[(4 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 5), ]$std.all,
#   tab_results_raw[(5 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 6), ]$std.all,
#   tab_results_raw[(6 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 7), ]$std.all,
#   tab_results_raw[(7 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 8), ]$std.all,
#   tab_results_raw[(8 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 9), ]$std.all,
#   tab_results_raw[(9 * length(vars_pers_fac) + 1) : (length(vars_pers_fac) * 10), ]$std.all
# )
# names(tab_results) <- vars_pri
# rownames(tab_results) <- vars_pers_fac
```

## Factors

Factor or dimensions.

```{r}
model_dim <- "
HEX_HOH =~ HEX_HOH_SIN + HEX_HOH_FAI + HEX_HOH_GRE + HEX_HOH_MOD
HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04
HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04
HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04

HEX_EMO =~ HEX_EMO_FEA + HEX_EMO_ANX + HEX_EMO_DEP + HEX_EMO_SEN
HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04
HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04
HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04
HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04

HEX_EXT =~ HEX_EXT_SSE + HEX_EXT_BOL + HEX_EXT_SOC + HEX_EXT_LIV
HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04
HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04
HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04
HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04

HEX_AGR =~ HEX_AGR_FOR + HEX_AGR_GEN + HEX_AGR_FLX + HEX_AGR_PAT
HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04
HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04
HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04
HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04

HEX_CNS =~ HEX_CNS_ORG + HEX_CNS_DIL + HEX_CNS_PER + HEX_CNS_PRU
HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04
HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04
HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04
HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04

HEX_OPN =~ HEX_OPN_AES + HEX_OPN_INQ + HEX_OPN_CRE + HEX_OPN_UNC
HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04
HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04
HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04
HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04

NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_CNC =~ NFP_CNC_01 + NFP_CNC_02 + NFP_CNC_03 + NFP_CNC_04
NFP_SPA =~ NFP_SPA_01 + NFP_SPA_02 + NFP_SPA_03 + NFP_SPA_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_INT =~ NFP_INT_01 + NFP_INT_02 + NFP_INT_03 + NFP_INT_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
"
fit_dim <- sem(model_dim, d, estimator = "ML", fixed.x = TRUE)
summary(fit_dim, standardized = TRUE)

# extract results
d_dim <- 
  standardizedsolution(fit_dim, ci = TRUE, level = .90) %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_pers_dim, rhs %in% vars_pri)

fig_dim <- ggplot(d_dim, aes(x = est.std, y = lhs)) +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper)) +
  facet_wrap(vars(rhs), nrow = 1)

# tab_results <- data.frame(
#   tab_results_raw[1 : length(vars_pers_dim), ]$std.all,
#   tab_results_raw[(1 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 2), ]$std.all,
#   tab_results_raw[(2 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 3), ]$std.all,
#   tab_results_raw[(3 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 4), ]$std.all,
#   tab_results_raw[(4 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 5), ]$std.all,
#   tab_results_raw[(5 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 6), ]$std.all,
#   tab_results_raw[(6 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 7), ]$std.all,
#   tab_results_raw[(7 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 8), ]$std.all,
#   tab_results_raw[(8 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 9), ]$std.all,
#   tab_results_raw[(9 * length(vars_pers_dim) + 1) : (length(vars_pers_dim) * 10), ]$std.all
# )
# names(tab_results) <- vars_pri
# rownames(tab_results) <- vars_pers_dim
```

## SES

```{r}
model_ses <- "
age =~ AGE
gender =~ GEN
ethnicity =~ ETH
relationship =~ REL_
college =~ COL_
income =~ INC
conservatism =~ CON
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_CNC =~ NFP_CNC_01 + NFP_CNC_02 + NFP_CNC_03 + NFP_CNC_04
NFP_SPA =~ NFP_SPA_01 + NFP_SPA_02 + NFP_SPA_03 + NFP_SPA_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_INT =~ NFP_INT_01 + NFP_INT_02 + NFP_INT_03 + NFP_INT_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
"
fit_ses <- sem(model_ses, d, estimator = "ML", fixed.x = TRUE)
summary(fit_ses, standardized = TRUE)

# extract results
d_ses <- 
  standardizedsolution(fit_ses, ci = TRUE, level = .90) %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_ses, rhs %in% vars_pri)

fig_ses <- ggplot(d_ses, aes(x = est.std, y = lhs)) +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper)) +
  facet_wrap(vars(rhs), nrow = 1)

# tab_results <- data.frame(
#   tab_results_raw[1 : length(vars_ses), ]$std.all,
#   tab_results_raw[(1 * length(vars_ses) + 1) : (length(vars_ses) * 2), ]$std.all,
#   tab_results_raw[(2 * length(vars_ses) + 1) : (length(vars_ses) * 3), ]$std.all,
#   tab_results_raw[(3 * length(vars_ses) + 1) : (length(vars_ses) * 4), ]$std.all,
#   tab_results_raw[(4 * length(vars_ses) + 1) : (length(vars_ses) * 5), ]$std.all,
#   tab_results_raw[(5 * length(vars_ses) + 1) : (length(vars_ses) * 6), ]$std.all,
#   tab_results_raw[(6 * length(vars_ses) + 1) : (length(vars_ses) * 7), ]$std.all,
#   tab_results_raw[(7 * length(vars_ses) + 1) : (length(vars_ses) * 8), ]$std.all,
#   tab_results_raw[(8 * length(vars_ses) + 1) : (length(vars_ses) * 9), ]$std.all,
#   tab_results_raw[(9 * length(vars_ses) + 1) : (length(vars_ses) * 10), ]$std.all
# )
# names(tab_results) <- vars_pri
# rownames(tab_results) <- vars_ses
```

```{r}
save.image("data/workspace.RData")
```