### 2023-03-01 items.rmd

---
title: "Items"
# output:
#   html_document:
#     toc: TRUE
#     toc_float:
#       collapsed: FALSE
#       toc_depth: 3
output: pdf_document
bibliography: 
  - "bibliography/references.bib"
---

All items were answered on a 7-point scale with the following options:

"To what extent do you agree or disagree with the following statements?"

(-3) Strongly Disagree, (-2) Disagree, (-1) Slightly Disagree, (0) Neutral, (1) Slightly Agree (2) Agree, (3) Strongly Agree

# Personality

@leePsychometricPropertiesHEXACO1002018

## Honesty Humility
### Sincerity

 1. If I want something from a person I dislike, I will act very nicely toward that person in order to get it.
 1. I wouldn't use flattery to get a raise or promotion at work, even if I thought it would succeed.
 1. If I want something from someone, I will laugh at that person's worst jokes.
 1. I wouldn't pretend to like someone just to get that person to do favors for me.

### Fairness

 1. If I knew that I could never get caught, I would be willing to steal a million dollars.
 1. I would be tempted to buy stolen property if I were financially tight.
 1. I would never accept a bribe, even if it were very large.
 1. I’d be tempted to use counterfeit money, if I were sure I could get away with it.

### Greed Avoidance

 1. Having a lot of money is not especially important to me.
 1. I would like to live in a very expensive, high-class neighborhood.
 1. I would like to be seen driving around in a very expensive car.
 1. I would get a lot of pleasure from owning expensive luxury goods.

### Modesty

 1. I am an ordinary person who is no better than others.
 1. I wouldn’t want people to treat me as though I were superior to them.
 1. I think that I am entitled to more respect than the average person is.
 1. I want people to know that I am an important person of high status.

## Emotionality
### Fearfulness

 1. I would feel afraid if I had to travel in bad weather conditions.
 1. I don’t mind doing jobs that involve dangerous work.
 1. When it comes to physical danger, I am very fearful.
 1. Even in an emergency I wouldn't feel like panicking.

### Anxiety

 1. I sometimes can't help worrying about little things.
 1. I worry a lot less than most people do.
 1. I rarely, if ever, have trouble sleeping due to stress or anxiety.
 1. I get very anxious when waiting to hear about an important decision.

### Dependence
 1. When I suffer from a painful experience, I need someone to make me feel comfortable.
 1. I can handle difficult situations without needing emotional support from anyone else.
 1. Whenever I feel worried about something, I want to share my concern with another person.
 1. I rarely discuss my problems with other people.

### Sentimentality

 1. I feel like crying when I see other people crying.
 1. When someone I know well is unhappy, I can almost feel that person's pain myself.
 1. I feel strong emotions when someone close to me is going away for a long time.
 1. I remain unemotional even in situations where most people get very sentimental.

## Extraversion
### Social Self-Esteem

 1. I feel reasonably satisfied with myself overall.
 1. I think that most people like some aspects of my personality.
 1. I feel that I am an unpopular person.
 1. Even in an emergency I wouldn't feel like panicking.

### Social Boldness

 1. I rarely express my opinions in group meetings.
 1. In social situations, I'm usually the one who makes the first move.
 1. When I'm in a group of people, I'm often the one who speaks on behalf of the group.
 1. I tend to feel quite self-conscious when speaking in front of a group of people.

### Sociability

 1. I avoid making "small talk" with people.
 1. I enjoy having lots of people around to talk with.
 1. I prefer jobs that involve active social interaction to those that involve working alone.
 1. The first thing that I always do in a new place is to make friends.

### Liveliness

 1. I am energetic nearly all the time.
 1. On most days, I feel cheerful and optimistic.
 1. People often tell me that I should try to cheer up.
 1. Most people are more upbeat and dynamic than I generally am.

## Agreeableness
### Forgiveness

 1. I rarely hold a grudge, even against people who have badly wronged me.
 1. My attitude toward people who have treated me badly is "forgive and forget".
 1. If someone has cheated me once, I will always feel suspicious of that person.
 1. I find it hard to fully forgive someone who has done something mean to me.

### Gentleness

 1. People sometimes tell me that I am too critical of others.
 1. I generally accept people’s faults without complaining about them.
 1. I tend to be lenient in judging other people.
 1. Even when people make a lot of mistakes, I rarely say anything negative.

### Flexibility

 1. People sometimes tell me that I'm too stubborn.
 1. I am usually quite flexible in my opinions when people disagree with me.
 1. When people tell me that I’m wrong, my first reaction is to argue with them.
 1. I find it hard to compromise with people when I really think I’m right.

### Patience

 1. People think of me as someone who has a quick temper.
 1. I rarely feel anger, even when people treat me quite badly.
 1. Most people tend to get angry more quickly than I do.
 1. I find it hard to keep my temper when people insult me.

## Conscientiousness
### Organization

 1. I clean my office or home quite frequently.
 1. I plan ahead and organize things, to avoid scrambling at the last minute.
 1. People often joke with me about the messiness of my room or desk.
 1. When working, I sometimes have difficulties due to being disorganized.

### Diligence

 1. When working, I often set ambitious goals for myself.
 1. I often push myself very hard when trying to achieve a goal.
 1. Often when I set a goal, I end up quitting without having reached it.
 1. I do only the minimum amount of work needed to get by.

### Perfectionism

 1. I often check my work over repeatedly to find any mistakes.
 1. When working on something, I don't pay much attention to small details.
 1. I always try to be accurate in my work, even at the expense of time.
 1. People often call me a perfectionist.

### Prudence

 1. I make decisions based on the feeling of the moment rather than on careful thought.
 1. I make a lot of mistakes because I don't think before I act.
 1. I don’t allow my impulses to govern my behavior.
 1. I prefer to do whatever comes to mind, rather than stick to a plan.

## Openness
### Aesthetic Appreciation

 1. I would be quite bored by a visit to an art gallery.
 1. I wouldn't spend my time reading a book of poetry.
 1. If I had the opportunity, I would like to attend a classical music concert.
 1. Sometimes I like to just watch the wind as it blows through the trees.

### Inquisitiveness

 1. I'm interested in learning about the history and politics of other countries.
 1. I enjoy looking at maps of different places.
 1. I would be very bored by a book about the history of science and technology.
 1. I’ve never really enjoyed looking through an encyclopedia.

### Creativeness

 1. I would like a job that requires following a routine rather than being creative.
 1. I would enjoy creating a work of art, such as a novel, a song, or a painting.
 1. People have often told me that I have a good imagination.
 1. I don't think of myself as the artistic or creative type.

### Unconventionality

 1. I think that paying attention to radical ideas is a waste of time.
 1. I like people who have unconventional views.
 1. I think of myself as a somewhat eccentric person.
 1. I find it boring to discuss philosophy.

## Altruism

 1. I have sympathy for people who are less fortunate than I am.
 1. I try to give generously to those in need.
 1. It wouldn’t bother me to harm someone I didn’t like.
 1. People see me as a hard-hearted person.

# Need for privacy
## Horizontal privacy
### Psychological

@frenerDevelopmentValidationNeed2021

 1.	I don’t like to talk about personal issues with others unless they do it first.
 2.	There are a lot of things about me that I don’t like to talk about with others.
 3.	I prefer not to share my feelings and inner thoughts with others.
 4.	I don't like it when others talk to me about their private issues. (new)

### Physical

@frenerDevelopmentValidationNeed2021

 1.	I don’t like it when strangers come too close to me.
 2.	I don’t like to stand in a dense crowd of people.
 3.	I don’t like to sit next to a stranger on a tram, bus, or plane.
 4.	I hate it when people enter my personal space uninvited. (new)

### Social

@dienlinWhoNeedsPrivacy2019

 1.	I don’t feel comfortable interacting with other people. [@harrisonDevelopmentDesirePrivacy1993]
 2.	I often want to be alone. (new)
 3.	I’m a fairly anti-social person. (new)
 4.	I prefer to work independently in most situations. (new)

## Vertical privacy
### Privacy from Government

@dienlinWhoNeedsPrivacy2019

 1.	I need government agencies to respect my privacy, even if that makes catching criminals harder.
 2.	I don’t want the government to gather data about me, even if that makes it harder to spend tax income effectively.
 3.	I don’t want government agencies to monitor my communication, even if that stops terrorist attacks.
 4.	I feel I need to protect my privacy from government agencies.

### Privacy from Companies

New, self-designed.

 1. I’m willing to pay more for products or services so that companies don’t have to sell my data.
 2. I’m happy to give up a little bit of privacy so that I can use certain apps or services for free.
 3. I don’t want companies to collect data about me, even if that makes their services worse.
 4. I feel I need to protect my privacy from companies.

## General privacy
### Informational

@frenerDevelopmentValidationNeed2021

 1.	I would prefer that little is known about me.
 2.	I don’t want my personal data to be publicly accessible.
 3.	It is important to me that records pertaining to me remain confidential. [@harrisonDevelopmentDesirePrivacy1993]
 4.	I prefer that others cannot easily find information about me on the Internet. (new)

### Privacy need anonymity

@dienlinWhoNeedsPrivacy2019

 1.	I would prefer to use a fake account on social network sites to preserve my privacy.
 2.	I feel I need to avoid places with video surveillance.
 3.	I prefer not to carry my driver’s license or ID with me all the time to preserve my privacy.
 4.	I need to be able to surf the Internet anonymously. (new)

### General Need for privacy

New, self-designed.

 1.	I need a lot of privacy.
 2.	Privacy is very important to me.
 3.	I think a lot about how I can protect my privacy.
 4.	I value privacy a lot.

# Privacy other
## Concerns

I'm very much concerned about my privacy in general.

## Behavior

I often take safety measures to protect my privacy.

# Sociodemographics

## Age

How old are you?

## Gender

How do you describe yourself?

 1. Man
 2. Woman
 3. Third gender / other

## Ethnicity

With what ethnicity do you identify with?

 1. White
 2. Other

## Relationship status

Are you currently in a partnership or relation with another person?

 1. No
 2. Yes

## College

Please indicate the highest degree you have completed

 1. No college degree
 2. College degree

## Income

About how much money do you have at your disposal each month?

 1. < \$1.000
 2. \$1.000 - \$4.000
 3. > \$4.000

## Politics

Where would you position yourself politically?

 - Extremely liberal
 - Liberal
 - Somewhat liberal
 - Neutral
 - Somewhat conservative
 - Conservative
 - Extremely conservative

## Literature