### 2023-03-01 items_need-for-privacy.Rmd

---
title: 'Items: Need for Privacy'
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      toc_depth: 3
bibliography: bibliography/references.bib
---

All items were answered on a 7-point scale with the following options:

"To what extent do you agree or disagree with the following statements?"

(-3) Strongly Disagree, (-2) Disagree, (-1) Slightly Disagree, (0) Neutral, (1) Slightly Agree (2) Agree, (3) Strongly Agree

# Need for privacy
## Horizontal privacy
### Psychological

@frenerDevelopmentValidationNeed2021

 1.	I don’t like to talk about personal issues with others unless they do it first.
 2.	There are a lot of things about me that I don’t like to talk about with others.
 3.	I prefer not to share my feelings and inner thoughts with others.
 4.	I do not like it when others talk to me about their private issues. (new)

### Physical

@frenerDevelopmentValidationNeed2021

 1.	I don’t like it when strangers come too close to me.
 2.	I don’t like to stand in a dense crowd of people.
 3.	I don’t like to sit next to a stranger on a tram, bus, or plane.
 4.	I hate it when people enter my personal space uninvited. (new)

### Social

@dienlinWhoNeedsPrivacy2019

 1.	I don’t feel comfortable interacting with other people. [@harrisonDevelopmentDesirePrivacy1993]
 2.	I often want to be alone. (new)
 3.	I’m a fairly anti-social person. (new)
 4.	I prefer to work independently in most situations. (new)

## Vertical privacy
### Privacy from Government

@dienlinWhoNeedsPrivacy2019

 1.	I need government agencies to respect my privacy, even if that makes catching criminals harder.
 2.	I don’t want the government to gather data about me, even if that makes it harder to spend tax income effectively.
 3.	I don’t want government agencies to monitor my communication, even if that prevents terrorist attacks.
 4.	I feel I need to protect my privacy from government agencies.

### Privacy from Companies

New, self-designed.

 1.	I’m willing to pay more for products or services so that companies don’t have to sell my data.
 2.	I’m happy to give up a little bit of privacy so that I can use certain apps or services for free.
 3.	I don’t want companies to gather information about me, even if that reduces the quality of their services.
 4.	I feel I need to protect my privacy from companies.

## General privacy
### Informational

@frenerDevelopmentValidationNeed2021

 1.	I would prefer that little is known about me.
 2.	I don’t want my personal data to be publicly accessible.
 3.	It is important to me that records pertaining to me remain confidential. [@harrisonDevelopmentDesirePrivacy1993]
 4.	I prefer that others cannot easily find information about me on the Internet. (new)

### Privacy need anonymity

@dienlinWhoNeedsPrivacy2019

 1.	I would prefer to use a fake account on social network sites to preserve my privacy.
 2.	I feel I need to avoid places with video surveillance.
 3.	I prefer not to carry my driver’s license or ID with me all the time to preserve my privacy.
 4.	I need to be able to surf the Internet anonymously. (new)

### General Need for privacy

New, self-designed.

1.	I need a lot of privacy.
2.	Privacy is very important to me.
3.	I think a lot about how I can protect my privacy.
4.	I value privacy a lot.

## Literature