### 2023-03-01 analyses.RmD
- added results to CFAs

---
title: "Who Needs Privacy? Exploring the relationship between personality and need for privacy"
subtitle: "Analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
    df_print: paged
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 90)
```

Here you can find the results of our study.

- At this point, we haven't yet collected data. 
- The results below use simulated data. 
- The data are random and don't make sense, but help to illustrate the overall approach and planned analyses.
- We'll later conduct further analyses not yet reported here, for example model-alterations given misfit.

# Set-Up

```{r r-setup, include=T, echo=T, results='hide'}
# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# define packages
packages <- c("corrplot", "devtools", "GGally", "ggplot2", "haven", "kableExtra", "knitr", 
              "lavaan", "magrittr", "mice", "MVN", "PerFit", "psych", "quanteda.textstats",
              "pwr", "semPlot", "semTools", "tidyverse", "td")

# load packages
lapply(packages, library, character.only = TRUE)

# readability function
readability = function(data){
  reading_estimates <- c("Flesch.Kincaid", "FOG", "Coleman.Liau.short", "Dale.Chall.PSK")
  textstat_readability(data, measure = reading_estimates) %>% 
    as.data.frame %>% 
    mutate(average = rowMeans(select(., -document)))
}
```

```{r data-wrangling}
d <- read_spss("data/Need+for+Privacy_February+27%2C+2023_13.45.sav")
d <- rename(d, COL = COL_, REL = REL_)

# define vars and labels
vars_pers_fac <- c("HEX_HOH_SIN", "HEX_HOH_FAI", "HEX_HOH_GRE", "HEX_HOH_MOD", 
                   "HEX_EMO_FEA", "HEX_EMO_ANX", "HEX_EMO_DEP", "HEX_EMO_SEN", 
                   "HEX_EXT_SSE", "HEX_EXT_BOL", "HEX_EXT_SOC", "HEX_EXT_LIV", 
                   "HEX_AGR_FOR", "HEX_AGR_GEN", "HEX_AGR_FLX", "HEX_AGR_PAT", 
                   "HEX_CNS_ORG", "HEX_CNS_DIL", "HEX_CNS_PER", "HEX_CNS_PRU", 
                   "HEX_OPN_AES", "HEX_OPN_INQ", "HEX_OPN_CRE", "HEX_OPN_UNC", 
                   "HEX_ALT")
vars_pers_fac_txt <- c("Sincerity", "Fairness", "Greed avoidance", "Modesty", 
                       "Fearfulness", "Anxiety", "Dependence", "Sentimentality", 
                       "Social self-esteem", "Social boldness", "Sociability", "Liveliness",
                       "Forgiveness", "Gentleness", "Flexibility", "Patience", 
                       "Organization", "Diligence", "Perfectionism", "Prudence", 
                       "Aesthetic appreciation", "Inquisitiveness", "Creativeness", "Unconventionality", 
                       "Altruism")

vars_pers_dim <- c("HEX_HOH", "HEX_EMO", "HEX_EXT", "HEX_AGR", "HEX_CNS", "HEX_OPN")
vars_pers_dim_txt <- c("Honesty humility", "Emotionality", "Extraversion", 
                       "Agreeableness", "Conscientiousness", "Openness")

vars_pri <- c("NFP_SOC", "NFP_PHY", "NFP_PSY",
              "NFP_GOV", "NFP_COM", 
              "NFP_INF", "NFP_ANO", "NFP_GEN")
vars_pri_txt <- c("Social", "Physical", "Psychological",
                  "Companies", "Governnment",
                  "Anonynmity", "Infornmational", "General")
vars_pri_txt_brk <- c("Social", "Physical", "Psycho-\nlogical",
                  "Companies", "Govern-\nment",
                  "Anony-\nmity", "Infor-\nmational", "General")
vars_pri_txt_abr <- c("Social", "Phys.", "Psych.",
                  "Comp.", "Gov.",
                  "Anonym.", "Inform.", "General")

vars_ses <- c("AGE", "GEN", "ETH", "REL", "COL", "INC", "CON")
vars_ses_txt <- c("Age", "Gender", "Ethnicity", "Relationship", "College", "Income", "Conservatism")
```

# Items

All items were answered on a 7-point scale with the following options:

"To what extent do you agree or disagree with the following statements?"

(-3) Strongly Disagree, (-2) Disagree, (-1) Slightly Disagree, (0) Neutral, (1) Slightly Agree (2) Agree, (3) Strongly Agree

## Personality
### Honesty Humility
#### Sincerity

 - If I want something from a person I dislike, I will act very nicely toward that person in order to get it.
 - I wouldn't use flattery to get a raise or promotion at work, even if I thought it would succeed.
 - If I want something from someone, I will laugh at that person's worst jokes.
 - I wouldn't pretend to like someone just to get that person to do favors for me.

```{r}
name <- "HEX_HOH_SIN"
model <- "HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Fairness

 - If I knew that I could never get caught, I would be willing to steal a million dollars.
 - I would be tempted to buy stolen property if I were financially tight.
 - I would never accept a bribe, even if it were very large.
 - I’d be tempted to use counterfeit money, if I were sure I could get away with it.

```{r}
name <- "HEX_HOH_FAI"
model <- "HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Greed Avoidance

 - Having a lot of money is not especially important to me.
 - I would like to live in a very expensive, high-class neighborhood.
 - I would like to be seen driving around in a very expensive car.
 - I would get a lot of pleasure from owning expensive luxury goods.

```{r}
name <- "HEY_HOH_GRE"
model <- "HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Modesty

 - I am an ordinary person who is no better than others.
 - I wouldn’t want people to treat me as though I were superior to them.
 - I think that I am entitled to more respect than the average person is.
 - I want people to know that I am an important person of high status.

```{r}
name <- "HEX_HOH_MOD"
model <- "HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Emotionality
#### Fearfulness

 - I would feel afraid if I had to travel in bad weather conditions.
 - I don’t mind doing jobs that involve dangerous work.
 - When it comes to physical danger, I am very fearful.
 - Even in an emergency I wouldn't feel like panicking.

```{r}
name <- "HEX_EMO_FEA"
model <- "HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Anxiety

 - I sometimes can't help worrying about little things.
 - I worry a lot less than most people do.
 - I rarely, if ever, have trouble sleeping due to stress or anxiety.
 - I get very anxious when waiting to hear about an important decision.

```{r}
name <- "HEX_EMO_ANX"
model <- "HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Dependence

 - When I suffer from a painful experience, I need someone to make me feel comfortable.
 - I can handle difficult situations without needing emotional support from anyone else.
 - Whenever I feel worried about something, I want to share my concern with another person.
 - I rarely discuss my problems with other people.

```{r}
name <- "HEX_EMO_DEP"
model <- "HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Sentimentality

 - I feel like crying when I see other people crying.
 - When someone I know well is unhappy, I can almost feel that person's pain myself.
 - I feel strong emotions when someone close to me is going away for a long time.
 - I remain unemotional even in situations where most people get very sentimental.

```{r}
name <- "HEX_EMO_SEN"
model <- "HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Extraversion
#### Social Self-Esteem

 - I feel reasonably satisfied with myself overall.
 - I think that most people like some aspects of my personality.
 - I feel that I am an unpopular person.
 - Even in an emergency I wouldn't feel like panicking.

```{r}
name <- "HEX_EXT_SSE"
model <- "HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Social Boldness

 - I rarely express my opinions in group meetings.
 - In social situations, I'm usually the one who makes the first move.
 - When I'm in a group of people, I'm often the one who speaks on behalf of the group.
 - I tend to feel quite self-conscious when speaking in front of a group of people.

```{r}
name <- "HEX_EXT_BOL"
model <- "HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Sociability

 - I avoid making "small talk" with people.
 - I enjoy having lots of people around to talk with.
 - I prefer jobs that involve active social interaction to those that involve working alone.
 - The first thing that I always do in a new place is to make friends.

```{r}
name <- "HEX_EXT_SOC"
model <- "HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Liveliness

 - I am energetic nearly all the time.
 - On most days, I feel cheerful and optimistic.
 - People often tell me that I should try to cheer up.
 - Most people are more upbeat and dynamic than I generally am.

```{r}
name <- "HEX_EXT_LIV"
model <- "HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Agreeableness
#### Forgiveness

 - I rarely hold a grudge, even against people who have badly wronged me.
 - My attitude toward people who have treated me badly is "forgive and forget".
 - If someone has cheated me once, I will always feel suspicious of that person.
 - I find it hard to fully forgive someone who has done something mean to me.

```{r}
name <- "HEX_AGR_FOR"
model <- "HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Gentleness

 - People sometimes tell me that I am too critical of others.
 - I generally accept people’s faults without complaining about them.
 - I tend to be lenient in judging other people.
 - Even when people make a lot of mistakes, I rarely say anything negative.

```{r}
name <- "HEX_AGR_GEN"
model <- "HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Flexibility

 - People sometimes tell me that I'm too stubborn.
 - I am usually quite flexible in my opinions when people disagree with me.
 - When people tell me that I’m wrong, my first reaction is to argue with them.
 - I find it hard to compromise with people when I really think I’m right.

```{r}
name <- "HEX_AGR_FLX"
model <- "HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Patience

 - People think of me as someone who has a quick temper.
 - I rarely feel anger, even when people treat me quite badly.
 - Most people tend to get angry more quickly than I do.
 - I find it hard to keep my temper when people insult me.

```{r}
name <- "HEX_AGR_PAT"
model <- "HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Conscientiousness
#### Organization

 - I clean my office or home quite frequently.
 - I plan ahead and organize things, to avoid scrambling at the last minute.
 - People often joke with me about the messiness of my room or desk.
 - When working, I sometimes have difficulties due to being disorganized.

```{r}
name <- "HEX_CNS_ORG"
model <- "HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Diligence

 - When working, I often set ambitious goals for myself.
 - I often push myself very hard when trying to achieve a goal.
 - Often when I set a goal, I end up quitting without having reached it.
 - I do only the minimum amount of work needed to get by.

```{r}
name <- "HEX_CNS_DIL"
model <- "HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Perfectionism

 - I often check my work over repeatedly to find any mistakes.
 - When working on something, I don't pay much attention to small details.
 - I always try to be accurate in my work, even at the expense of time.
 - People often call me a perfectionist.

```{r}
name <- "HEX_CNS_PER"
model <- "HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Prudence

 - I make decisions based on the feeling of the moment rather than on careful thought.
 - I make a lot of mistakes because I don't think before I act.
 - I don’t allow my impulses to govern my behavior.
 - I prefer to do whatever comes to mind, rather than stick to a plan.

```{r}
name <- "HEX_CNS_PRU"
model <- "HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Openness
#### Aesthetic Appreciation

 - I would be quite bored by a visit to an art gallery.
 - I wouldn't spend my time reading a book of poetry.
 - If I had the opportunity, I would like to attend a classical music concert.
 - Sometimes I like to just watch the wind as it blows through the trees.

```{r}
name <- "HEX_OPN_AEX"
model <- "HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Inquisitiveness

 - I'm interested in learning about the history and politics of other countries.
 - I enjoy looking at maps of different places.
 - I would be very bored by a book about the history of science and technology.
 - I’ve never really enjoyed looking through an encyclopedia.

```{r}
name <- "HEX_OPN_INQ"
model <- "HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Creativeness

 - I would like a job that requires following a routine rather than being creative.
 - I would enjoy creating a work of art, such as a novel, a song, or a painting.
 - People have often told me that I have a good imagination.
 - I don't think of myself as the artistic or creative type.

```{r}
name <- "HEX_OPN_CRE"
model <- "HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

#### Unconventionality

 - I think that paying attention to radical ideas is a waste of time.
 - I like people who have unconventional views.
 - I think of myself as a somewhat eccentric person.
 - I find it boring to discuss philosophy.

```{r}
name <- "HEX_OPN_UNC"
model <- "HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Altruism

 - I have sympathy for people who are less fortunate than I am.
 - I try to give generously to those in need.
 - It wouldn’t bother me to harm someone I didn’t like.
 - People see me as a hard-hearted person.

```{r}
name <- "HEX_ALT"
model <- "HEX_ALT =~ HEX_ALT_01 + HEX_ALT_02 + HEX_ALT_03 + HEX_ALT_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

## Need for privacy
### Psychological

 1.	I don’t like to talk about personal issues with others unless they do it first.
 2.	There are a lot of things about me that I don’t like to talk about with others.
 3.	I prefer not to share my feelings and inner thoughts with others.
 4.	I don't like it when others talk to me about their private issues.

```{r}
name <- "NFP_PSY"
model <- "NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Physical

 1.	I don’t like it when strangers come too close to me.
 2.	I don’t like to stand in a dense crowd of people.
 3.	I don’t like to sit next to a stranger on a tram, bus, or plane.
 4.	I hate it when people enter my personal space uninvited.

```{r}
name <- "NFP_PHY"
model <- "NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Social

 1.	I don’t feel comfortable interacting with other people.
 2.	I often want to be alone.
 3.	I’m a fairly anti-social person.
 4.	I prefer to work independently in most situations.

```{r}
name <- "NFP_SOC"
model <- "NFP_SOC =~ NFP_SOC_01 + NFP_SOC_02 + NFP_SOC_03 + NFP_SOC_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Privacy from Government

 1.	I need government agencies to respect my privacy, even if that makes catching criminals harder.
 2.	I don’t want the government to gather data about me, even if that makes it harder to spend tax income effectively.
 3.	I don’t want government agencies to monitor my communication, even if that stops terrorist attacks.
 4.	I feel I need to protect my privacy from government agencies.

```{r}
name <- "NFP_GOV"
model <- "NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Privacy from Companies

 1. I’m willing to pay more for products or services so that companies don’t have to sell my data.
 2. I’m happy to give up a little bit of privacy so that I can use certain apps or services for free.
 3. I don’t want companies to collect data about me, even if that makes their services worse.
 4. I feel I need to protect my privacy from companies.

```{r}
name <- "NFP_COM"
model <- "NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Informational

 1.	I would prefer that little is known about me.
 2.	I don’t want my personal data to be publicly accessible.
 3.	It is important to me that records pertaining to me remain confidential.
 4.	I prefer that others cannot easily find information about me on the Internet.

```{r}
name <- "NFP_INF"
model <- "NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### Privacy need anonymity

 1.	I would prefer to use a fake account on social network sites to preserve my privacy.
 2.	I feel I need to avoid places with video surveillance.
 3.	I prefer not to carry my driver’s license or ID with me all the time to preserve my privacy.
 4.	I need to be able to surf the Internet anonymously.

```{r}
name <- "NFP_ANO"
model <- "NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

### General Need for privacy

 1.	I need a lot of privacy.
 2.	Privacy is very important to me.
 3.	I think a lot about how I can protect my privacy.
 4.	I value privacy a lot.

```{r}
name <- "NFP_GEN"
model <- "NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04"
assign(paste("fit", name, sep = "_"), cfa(model, d, missing = "FIML"))
summary(get(paste0("fit_", name)), fit = TRUE, standardized = TRUE)
```

# Power Analyses

To estimate sample size, we conducted a rudimentary power analysis. That is, we planned for a two-sided correlation of _r_ = .1. We adopted a balanced alpha and beta error approach. Because we were in "exploration mode", we aimed for a high statistical power and hence relaxed alpha. Ideally, we would have wanted to set an error rate of 5%; however, this was outside of our budget. Instead, we opted for an error rate of 10%.

```{r}
sesoi <- .1

# a priori power analyses
pwr.r.test(r = sesoi, sig.level = .05, power = .95)
power_n_95 <- pwr.r.test(r = sesoi, sig.level = .05, power = .95)$n
```

A 5% error rate would imply collecting `r power_n_95` participants.

# Results
## SES

```{r}
model_ses <- "
# SES
Age =~ AGE
Gender =~ GEN
Ethnicity =~ ETH
Relationship =~ REL
College =~ COL
Income =~ INC
Conservatism =~ CON

# Privacy
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_SOC =~ NFP_SOC_01 + NFP_SOC_02 + NFP_SOC_03 + NFP_SOC_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
"
fit_ses <- sem(model_ses, d, estimator = "ML", fixed.x = TRUE)
summary(fit_ses, standardized = TRUE)

# get standardized data
d_ses <- 
  standardizedsolution(fit_ses)
```

```{r}
# make table
tab_ses <- 
  d_ses %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_ses_txt, rhs %in% vars_pri) %>% 
  select("Sociodemographics" = lhs, "Need for privacy" = rhs, est.std) %>% 
  mutate(`Need for privacy` = factor(`Need for privacy`, levels = vars_pri, labels = vars_pri_txt)) %>% 
  pivot_wider(id_cols = "Sociodemographics", names_from = "Need for privacy", values_from = "est.std") %>% 
  as.data.frame()
tab_ses
```

## Dimensions

Factor or dimensions.

```{r}
model_dim <- "
# Personality Factors
HEX_HOH =~ HEX_HOH_SIN + HEX_HOH_FAI + HEX_HOH_GRE + HEX_HOH_MOD
HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04
HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04
HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04

HEX_EMO =~ HEX_EMO_FEA + HEX_EMO_ANX + HEX_EMO_DEP + HEX_EMO_SEN
HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04
HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04
HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04
HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04

HEX_EXT =~ HEX_EXT_SSE + HEX_EXT_BOL + HEX_EXT_SOC + HEX_EXT_LIV
HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04
HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04
HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04
HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04

HEX_AGR =~ HEX_AGR_FOR + HEX_AGR_GEN + HEX_AGR_FLX + HEX_AGR_PAT
HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04
HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04
HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04
HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04

HEX_CNS =~ HEX_CNS_ORG + HEX_CNS_DIL + HEX_CNS_PER + HEX_CNS_PRU
HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04
HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04
HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04
HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04

HEX_OPN =~ HEX_OPN_AES + HEX_OPN_INQ + HEX_OPN_CRE + HEX_OPN_UNC
HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04
HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04
HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04
HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04

# Privacy
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_SOC =~ NFP_SOC_01 + NFP_SOC_02 + NFP_SOC_03 + NFP_SOC_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
"
fit_dim <- sem(model_dim, d, estimator = "ML", fixed.x = TRUE)
summary(fit_dim, standardized = TRUE)

# get standardized data
d_dim <- 
  standardizedsolution(fit_dim)
```

```{r}
# make table
tab_dim <- 
  d_dim %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_pers_dim, rhs %in% vars_pri) %>% 
  select("Personality dimension" = lhs, "Need for privacy" = rhs, est.std) %>% 
  mutate(`Need for privacy` = factor(`Need for privacy`, levels = vars_pri, labels = vars_pri_txt)) %>% 
  pivot_wider(id_cols = "Personality dimension", names_from = "Need for privacy", values_from = "est.std") %>% 
  mutate(`Personality dimension` = factor(`Personality dimension`, levels = vars_pers_dim, labels = vars_pers_dim_txt)) %>% 
  as.data.frame()
tab_dim
```

## Facets

```{r}
model_fac <- "
# Personality
HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04
HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04
HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04
HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04
HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04
HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04
HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04
HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04
HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04
HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04
HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04
HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04
HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04
HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04
HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04
HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04
HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04
HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04
HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04
HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04
HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04
HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04
HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04
HEX_ALT =~ HEX_ALT_01 + HEX_ALT_02 + HEX_ALT_03 + HEX_ALT_04

# Privacy
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_SOC =~ NFP_SOC_01 + NFP_SOC_02 + NFP_SOC_03 + NFP_SOC_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04
"
fit_fac <- sem(model_fac, d, estimator = "ML", fixed.x = TRUE)
summary(fit_fac, standardized = TRUE)

# get standardized data
d_fac <- 
  standardizedsolution(fit_fac)
```

```{r}
# make table
tab_fac <- 
  d_fac %>% 
  as.data.frame() %>% 
  filter(op == "~~", lhs %in% vars_pers_fac, rhs %in% vars_pri) %>% 
  select("Personality facet" = lhs, "Need for privacy" = rhs, est.std) %>% 
  mutate(`Need for privacy` = factor(`Need for privacy`, levels = vars_pri, labels = vars_pri_txt)) %>% 
  pivot_wider(id_cols = "Personality facet", names_from = "Need for privacy", values_from = "est.std") %>% 
  mutate(`Personality facet` = factor(`Personality facet`, levels = vars_pers_fac, labels = vars_pers_fac_txt)) %>% 
  as.data.frame
tab_fac
```

## Regression

```{r}
model_reg <- "
# Personality Factors
HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
HEX_HOH_FAI =~ HEX_HOH_FAI_01 + HEX_HOH_FAI_02 + HEX_HOH_FAI_03 + HEX_HOH_FAI_04
HEX_HOH_GRE =~ HEX_HOH_GRE_01 + HEX_HOH_GRE_02 + HEX_HOH_GRE_03 + HEX_HOH_GRE_04
HEX_HOH_MOD =~ HEX_HOH_MOD_01 + HEX_HOH_MOD_02 + HEX_HOH_MOD_03 + HEX_HOH_MOD_04

HEX_EMO_FEA =~ HEX_EMO_FEA_01 + HEX_EMO_FEA_02 + HEX_EMO_FEA_03 + HEX_EMO_FEA_04
HEX_EMO_ANX =~ HEX_EMO_ANX_01 + HEX_EMO_ANX_02 + HEX_EMO_ANX_03 + HEX_EMO_ANX_04
HEX_EMO_DEP =~ HEX_EMO_DEP_01 + HEX_EMO_DEP_02 + HEX_EMO_DEP_03 + HEX_EMO_DEP_04
HEX_EMO_SEN =~ HEX_EMO_SEN_01 + HEX_EMO_SEN_02 + HEX_EMO_SEN_03 + HEX_EMO_SEN_04

HEX_EXT_SSE =~ HEX_EXT_SSE_01 + HEX_EXT_SSE_02 + HEX_EXT_SSE_03 + HEX_EXT_SSE_04
HEX_EXT_BOL =~ HEX_EXT_BOL_01 + HEX_EXT_BOL_02 + HEX_EXT_BOL_03 + HEX_EXT_BOL_04
HEX_EXT_SOC =~ HEX_EXT_SOC_01 + HEX_EXT_SOC_02 + HEX_EXT_SOC_03 + HEX_EXT_SOC_04
HEX_EXT_LIV =~ HEX_EXT_LIV_01 + HEX_EXT_LIV_02 + HEX_EXT_LIV_03 + HEX_EXT_LIV_04

HEX_AGR_FOR =~ HEX_AGR_FOR_01 + HEX_AGR_FOR_02 + HEX_AGR_FOR_03 + HEX_AGR_FOR_04
HEX_AGR_GEN =~ HEX_AGR_GEN_01 + HEX_AGR_GEN_02 + HEX_AGR_GEN_03 + HEX_AGR_GEN_04
HEX_AGR_FLX =~ HEX_AGR_FLX_01 + HEX_AGR_FLX_02 + HEX_AGR_FLX_03 + HEX_AGR_FLX_04
HEX_AGR_PAT =~ HEX_AGR_PAT_01 + HEX_AGR_PAT_02 + HEX_AGR_PAT_03 + HEX_AGR_PAT_04

HEX_CNS_ORG =~ HEX_CNS_ORG_01 + HEX_CNS_ORG_02 + HEX_CNS_ORG_03 + HEX_CNS_ORG_04
HEX_CNS_DIL =~ HEX_CNS_DIL_01 + HEX_CNS_DIL_02 + HEX_CNS_DIL_03 + HEX_CNS_DIL_04
HEX_CNS_PER =~ HEX_CNS_PER_01 + HEX_CNS_PER_02 + HEX_CNS_PER_03 + HEX_CNS_PER_04
HEX_CNS_PRU =~ HEX_CNS_PRU_01 + HEX_CNS_PRU_02 + HEX_CNS_PRU_03 + HEX_CNS_PRU_04

HEX_OPN_AES =~ HEX_OPN_AES_01 + HEX_OPN_AES_02 + HEX_OPN_AES_03 + HEX_OPN_AES_04
HEX_OPN_INQ =~ HEX_OPN_INQ_01 + HEX_OPN_INQ_02 + HEX_OPN_INQ_03 + HEX_OPN_INQ_04
HEX_OPN_CRE =~ HEX_OPN_CRE_01 + HEX_OPN_CRE_02 + HEX_OPN_CRE_03 + HEX_OPN_CRE_04
HEX_OPN_UNC =~ HEX_OPN_UNC_01 + HEX_OPN_UNC_02 + HEX_OPN_UNC_03 + HEX_OPN_UNC_04

# Privacy
NFP_INF =~ NFP_INF_01 + NFP_INF_02 + NFP_INF_03 + NFP_INF_04
NFP_PSY =~ NFP_PSY_01 + NFP_PSY_02 + NFP_PSY_03 + NFP_PSY_04
NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
NFP_SOC =~ NFP_SOC_01 + NFP_SOC_02 + NFP_SOC_03 + NFP_SOC_04
NFP_GOV =~ NFP_GOV_01 + NFP_GOV_02 + NFP_GOV_03 + NFP_GOV_04
NFP_COM =~ NFP_COM_01 + NFP_COM_02 + NFP_COM_03 + NFP_COM_04
NFP_ANO =~ NFP_ANO_01 + NFP_ANO_02 + NFP_ANO_03 + NFP_ANO_04
NFP_GEN =~ NFP_GEN_01 + NFP_GEN_02 + NFP_GEN_03 + NFP_GEN_04

NFP_INF + NFP_PSY + NFP_PHY + NFP_SOC + NFP_GOV + NFP_COM + NFP_ANO + NFP_GEN ~ HEX_HOH_SIN + HEX_HOH_FAI + HEX_HOH_GRE + HEX_HOH_MOD + HEX_EMO_FEA + HEX_EMO_ANX + HEX_EMO_DEP + HEX_EMO_SEN + HEX_EXT_SSE + HEX_EXT_BOL + HEX_EXT_SOC + HEX_EXT_LIV + HEX_AGR_FOR + HEX_AGR_GEN + HEX_AGR_FLX + HEX_AGR_PAT + HEX_CNS_ORG + HEX_CNS_DIL + HEX_CNS_PER + HEX_CNS_PRU + HEX_OPN_AES + HEX_OPN_INQ + HEX_OPN_CRE + HEX_OPN_UNC + AGE + GEN + ETH + REL + COL + INC + CON
"
fit_reg <- sem(model_reg, d, estimator = "ML", fixed.x = TRUE)
summary(fit_reg, standardized = TRUE)

# get standardized data
d_reg <- 
  standardizedsolution(fit_reg, ci = TRUE)
```

```{r fig.width=10, fig.height=9}
# make table
tab_reg <- 
  d_reg %>% 
  as.data.frame() %>% 
  filter(op == "~", lhs %in% vars_pri) %>% 
  select("Need for privacy" = lhs, "Predictor" = rhs, Estimate = est.std, ci.lower, ci.upper) %>% 
  mutate(Predictor = factor(Predictor, levels = c(vars_pers_fac, vars_ses), labels = c(vars_pers_fac_txt, vars_ses_txt)),
         `Need for privacy` = factor(`Need for privacy`, levels = c(vars_pri), labels = c(vars_pri_txt)),
         Predictor_cat = c(vars_pers_dim_txt %>% rep(each = 4), "Sociodemographics" %>% rep(7)) %>% rep(length(vars_pri_txt)))

reg_fig <- ggplot(tab_reg, aes(x = Estimate, y = Predictor)) + 
  coord_cartesian(xlim = c(-.5, .5)) +
  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
  geom_vline(xintercept = -.1, lwd = .75, colour = "darkgrey", linetype = "dashed") + 
  geom_vline(xintercept = .1, lwd = .75, colour = "darkgrey", linetype = "dashed") + 
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper),
                lwd = .75, height = .2, position = position_dodge(-.7)) +
  geom_point(size = 1.5, position = position_dodge(-.7)) +
  scale_x_continuous(breaks = c(-.3, 0, .3)) + 
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    # panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    axis.text.y = element_text(hjust = 1),
    legend.position = "none"
    ) + 
    facet_grid(Predictor_cat ~ `Need for privacy`, scales = "free_y")
suppressWarnings(print(reg_fig, warning = F))

ggsave("figures/fig_reg.png", width = 9, height = 10)
```

```{r cache=F}
save.image("data/workspace.RData")
```