### 2024-02-08 index.Rmd
- Rendered website

---
title: "Who needs privacy?"
---

On this website, you can find the companion material for the paper _Who needs privacy? Exploring the relations between need for privacy and personality_. The paper is currently submitted to the journal Collabra as registered report. The report is based on an [older paper](https://osf.io/m23bn) that was already submitted to Collabra but rejected. This paper now addresses the [feedback from the original submission](reviews.pdf) and considers the original paper a prestudy.

- You can either read the paper [in your browser](proposal.html) or download it as [pdf](proposal.pdf). 
- You can also find the detailed code and summaries of the [analyses](analyses.html) reported in the paper

This website is produced directly from the project's [github repository](https://github.com/tdienlin/who_needs_privacy_rr). You can download the material and the data, file an issue, or submit a pull request.