### 2024-04-12 power_analyses.Rmd

---
title: "Power Analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
    df_print: paged
---

```{r knitr-setup, include=F}
set.seed(170819)

knitr::opts_chunk$set(
  warning=F, 
  echo=T, 
  error=F, 
  message=F, 
  cache = T, 
  tidy=T, 
  dev = 'pdf', 
  cache.extra = knitr::rand_seed,
  fig.height = 6,
  fig.width = 8
  )

options(
  digits = 3, 
  width = 150,
  max.print = 100000000
  )
```

```{r results=F, }
library(lavaan)
library(semTools)
library(simsem)
library(pwr)
```

# Power Analyses
## Rudimentary

To estimate sample size, we conducted a rudimentary power analysis. That is, we planned for a two-sided correlation of _r_ = .1. We adopted a balanced alpha and beta error approach. Because we were in "exploration mode", we aimed for a high statistical power and hence relaxed alpha. Ideally, we would have wanted to set an error rate of 5%; however, this was outside of our budget. Instead, we opted for an error rate of 10%.

```{r}
sesoi <- .1

# a priori power analyses
pwr.r.test(r = sesoi, sig.level = .05, power = .95)
power_n_95 <- pwr.r.test(r = sesoi, sig.level = .05, power = .95)$n
```

A 5% error rate would imply collecting `r power_n_95` participants.

## Detailed: Simple

```{r}
population <- 
  "
  # Fix means to 0
  HEX_HOH_SIN_01 ~ 0*1
  HEX_HOH_SIN_02 ~ 0*1
  HEX_HOH_SIN_03 ~ 0*1
  HEX_HOH_SIN_04 ~ 0*1
  NFP_PHY_01 ~ 0*1
  NFP_PHY_02 ~ 0*1
  NFP_PHY_03 ~ 0*1
  NFP_PHY_04 ~ 0*1

  # Fix variances
  HEX_HOH_SIN_01 ~~ 1 * HEX_HOH_SIN_01
  HEX_HOH_SIN_02 ~~ 1 * HEX_HOH_SIN_02
  HEX_HOH_SIN_03 ~~ 1 * HEX_HOH_SIN_03
  HEX_HOH_SIN_04 ~~ 1 * HEX_HOH_SIN_04
  NFP_PHY_01 ~~ 1 * NFP_PHY_01
  NFP_PHY_02 ~~ 1 * NFP_PHY_02
  NFP_PHY_03 ~~ 1 * NFP_PHY_03
  NFP_PHY_04 ~~ 1 * NFP_PHY_04

  # Define latent vars
  HEX_HOH_SIN =~ .9*HEX_HOH_SIN_01 + .9*HEX_HOH_SIN_02 + .9*HEX_HOH_SIN_03 + .9*HEX_HOH_SIN_04
  NFP_PHY =~ .9*NFP_PHY_01 + .9*NFP_PHY_02 + .9*NFP_PHY_03 + .9*NFP_PHY_04

  # Covariances
  HEX_HOH_SIN ~~ .1*NFP_PHY
  "
sample <- "
  HEX_HOH_SIN =~ HEX_HOH_SIN_01 + HEX_HOH_SIN_02 + HEX_HOH_SIN_03 + HEX_HOH_SIN_04
  NFP_PHY =~ NFP_PHY_01 + NFP_PHY_02 + NFP_PHY_03 + NFP_PHY_04
"

reliability <- simsem::sim(
  # nRep = NULL, 
  model = sample, 
#  n = 500,
  n = seq(500, 5000, 50), 
  generate = population, 
  # std.lv = TRUE, 
  lavaanfun = "sem"
  )

summary(reliability)
power_tab <- getPower(reliability)
power_80 <- findPower(power_tab, "N", 0.90)
power_80["HEX_HOH_SIN~~NFP_PHY"]
```

```{r}
save.image("data/workspace_1.RData")
```