### 2024-04-12 questionnaire.rmd

---
title: "Questionnaire"
subtitle: "Data-set: Need for Privacy and Personality"
author: "Tobias Dienlin & Miriam Metzger"
date: "2024-04-05"
# output:
#   html_document:
#     toc: TRUE
#     toc_float:
#       collapsed: FALSE
#       toc_depth: 3
output:
  pdf_document
bibliography: 
  - "bibliography/references.bib"
---

# Suggested citation

Dienlin, Tobias; Metzger, Miriam, 2024, "Need for Privacy and Personality", https://doi.org/10.11587/IC66GC, AUSSDA.

This document is licenced under a [Creative Commons Attribution 4.0 International Licence](https://creativecommons.org/licenses/by/4.0/).

```{r, out.width = "200px", fig.align='center', echo=F}
knitr::include_graphics("../Data/license.png")
```